---
title: "MUSA508_FINAL_SFPARK"
date: "December 10, 2021"
editor_options: 
  chunk_output_type: console
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
### PROMPT

**Project option 7: Forecasting parking demand What drives parking demand
(revenues)?** 
If you knew, could you create a tool that would predict parking demand over
time and space. Using San Francisco open parking data (locations), forecast parking demand as a function of built environment, neighborhood and road characteristics. Lots of app use cases possible depending if the user is public or private sector. There are 105m rows here – so you’ll have to use the API to download a small slice of the data.

## Use Case & Motivation
*Motivate the analysis – “What is the use case; why would someone want to
replicate your analysis and why would they use this approach?”*

## Method
*What is the spatial or space/time process?*

``` {r intro}
library(tidyverse)
library(sf)
library(lubridate)
library(tigris)
library(tidycensus)
library(viridis)
library(riem)
library(gridExtra)
library(knitr)
library(kableExtra)
library(RSocrata)
library(glue)
```

# 1. Data
*Describe the data you used.*

## 1.1 Import
``` {r data_import_park}
# https://data.sfgov.org/Transportation/SFMTA-Parking-Meter-Detailed-Revenue-Transactions/imvp-dq3v

park.cols = c(
'TRANSMISSION_DATETIME'='id.park',
#A system-generated unique identifier for the transaction record

'POST_ID'='id.meter',
#The identifier of the meter this transaction is related to. See the related meters dataset including locations https://data.sfgov.org/d/8vzz-qzz9

'STREET_BLOCK'='id.block',
#The name of the street block the meter sits on

'PAYMENT_TYPE'='payment.type',
#How the customer paid, one of the following: CASH, CREDIT CARD, SMART CARD

'SESSION_START_DT'='date.start',
#The date and time of the start of the meter session for this transaction by the customer

'SESSION_END_DT'='date.end',
#The date and time of the end of the meter session for this transaction by the customer

'METER_EVENT_TYPE'='event',
#Event type of the meter, one of the following: NS = New Session; AT = Addtional Time; SC = Status Change

'GROSS_PAID_AMT'='payment'
#The amount paid by the customer for this meter transaction

)

API.park.base = "https://data.sfgov.org/resource/imvp-dq3v.json"

API.park.fields = paste(c(names(park.cols)), collapse=', ')
API.park.fields = glue("$select={API.park.fields}")

date_field = 'SESSION_START_DT'
start_date = "2019-01-01T00:00:00.000"
end_date = "2020-01-01T00:00:00.000"
API.park.dates = glue("$where={date_field} between '{start_date}' and '{end_date}'")

API.park.add = paste(
    API.park.dates,
    API.park.fields,
    sep='?'
  ) %>% URLencode(.)
API.park.url = paste(
    API.park.base,
    API.park.add,
    sep='?'
  ) %>% URLencode(.)
park.data = read.socrata(API.park.url)
```

``` {r data_import_meter}

meter.cols = c(
'OBJECTID'='        id.object',
#
'PARKING_SPACE_ID'= 'id.park',
# Parking Space ID, connected to the physical location rather than the meter
'POST_ID'='id.meter',
# Unique identifier of meter
'MS_PAY_STATION_ID'='id.paystation',
#
'MS_SPACE_NUM'='MS_SPACE_NUM',
#
'SENSOR_FLAG'='SENSOR_FLAG',
# If there was an active occupancy sensor located at the space during the SFpark pilot (until Dec 2013)--N- No,Y - Yes
'ON_OFFSTREET_TYPE'='ON_OFFSTREET_TYPE',
# Whether the space is on the street or off‐street, e.g. metered parking lot--OFF Off‐street,ON On‐street
'OSP_ID'='OSP_ID',
# Unique identifier of SFMTA‐owned off‐street parking space--Null - Meter located on‐street,890 - Pier 48 Lot,891 - Pier 52 Lot,892 - Pier 1/2 Motorcycle Lot,901 - 24th and Capp Lot,902 - California and Steiner Lot,903 - 8th and Clement Lot,904 - 9th and Clement Lot,905 - Castro Theater Lot,906 - 18th and Collingwood Lot,907 - Mission and Norton Lot,908 - 21st and Geary Lot,909 - 18th and Geary Lot,910 - 20th and Irving Lot,911 - 8th and Irving Lot,913 - 7th and Irving Lot,914 - Junipero Serra and Ocean Lot,915 - 19th and Ocean Lot,916 - Pierce Street Garage,918 - 24th and Noe Lot,919 - Felton and San Bruno Lot,920 - SF General Hospital Lot,922 - West Portal Lot,923 - Claremont and Ulloa Lot,924 - Phelan Loop Lot
'JURISDICTION'='JURISDICTION',
# Agency which manages the meter--PORT - Port of San Francisco,SFMTA - San Francisco Municipal Transportation Agency
'PM_DISTRICT_ID'='PM_DISTRICT_ID',
# Parking Management District ID
'BLOCKFACE_ID'='BLOCKFACE_ID',
# Blockface (side of street) ID
'ACTIVE_METER_FLAG'='ACTIVE_METER_FLAG',
# Status of meter--M - Active,T - Temporarily inactive
'REASON_CODE'='REASON_CODE',
#
'SMART_METER_FLAG'='SMART_METER_FLAG',
#
'METER_TYPE'='METER_TYPE',
# How many spaces the meter manages-SS - Single‐space,MS - Multi‐space
'METER_VENDOR'='METER_VENDOR',
# Meter vendor
'METER_MODEL'='METER_MODEL',
# Meter model
'CAP_COLOR'='CAP_COLOR',
# Cap color describes the use and restrictions of the meter--Black - Motorcycle parking,Brown - Tour bus parking,Green - Short term parking,Grey - General metered parking,Purple - Boat trailer parking,Red - Six wheeled commercial vehicle parking,Yellow - Commercial vehicle parking
'PCO_BEAT'='PCO_BEAT',
#
'OLD_RATE_AREA'='OLD_RATE_AREA',
# Hourly rate zone of the meter; some variations within each zone--Area 1 - Downtown primarily; 3.50,Area 2 -Surrounding downtown primarily; 3.00,Area 3 -Residential neighborhood primarily; 2.00,Area 5 -SFpark area variable rates ranging from; 0.25 to; 6.00,MC1 -Motorcycle in Area 1 primarily; 0.70,MC2 -Motorcycle in Area 2 primarily; 0.60,MC3 -Motorcycle in Area 3 primarily; 0.40,MC5 -Motorcycle in SFpark area variable rates ranging from; 0.25 to; 6.00,Port 1 -Port‐owned at Fisherman’s Wharf; 2.50,Port 2 -Port‐owned at Fisherman’s Wharf; 2.50,Port 3 -Port‐owned at North Embarcadero; 2.00,Port 4 -Port‐owned at North Embarcadero; 2.00,Port 5 -Port‐owned at Downtown; 3.00,Port 6 -Port‐owned at Downtown; 3.00,Port 7 -Port‐owned at Downtown; 3.00,Port 8 -Port‐owned at Downtown; 3.00,Port 9 -Port‐owned at South Embarcadero; 1.00,Port 10 -Port‐owned at South Embarcadero; 1.00,Port 11 -Port‐owned at South Embarcadero; 1.00,Port 12 -Port‐owned at South Embarcadero; 1.00,PortMC1 -Port‐owned motorcycle rate; 0.25,PortMC2 -Port‐owned motorcycle rate at Ferry Building; 0.50,Tour Bus -; 9.00 for two hours
'STREET_ID'='STREET_ID',
# Street identifier
'STREET_NAME'='STREET_NAME',
# Street name
'STREET_NUM'='STREET_NUM',
# Approximate street  of meter
'PARITY_DIGIT_POSITION'='PARITY_DIGIT_POSITION',
#
'STREET_SEG_CTRLN_ID'='STREET_SEG_CTRLN_ID',
# Street segment centerline identifier
'ORIENTATION'='ORIENTATION',
#
'LONGITUDE'='LONGITUDE',
# Longitude of meter
'LATITUDE'='LATITUDE',
# Latitude of meter
'LEGISLATION_REF'='LEGISLATION_REF',
#
'LEGISLATION_DT'='LEGISLATION_DT',
#
'WORK_ORDER'='WORK_ORDER',
# Work order
'COMMENTS'='COMMENTS',
#
'COLLECTION_ROUTE'='COLLECTION_ROUTE',
#
'COLLECTION_SUBROUTE'='COLLECTION_SUBROUTE',
#
'PMR_ROUTE'='PMR_ROUTE',
#
'NFC_KEY'='NFC_KEY',
#
'SPT_CODE'='SPT_CODE',
#
'COLLECTION_ROUTE_DESC'='COLLECTION_ROUTE_DESC',
#
'COLLECTION_SUBROUTE_DESC'='COLLECTION_SUBROUTE_DESC',
#
'shape'='shape'
#
  )


API.meter.url = "https://data.sfgov.org/resource/imvp-dq3v.json"
meter.data = read.socrata(API.meter.url)
```

## Meter Filters

occupied rate = 80%

find threshold:
 - appropriate usage Rate by hour

Time Limit = 240 min (9am to 10pm)

``` {r data}
```


# 2. Exploratory Data Analysis
*Describe your exploratory analysis using maps and plots.*

``` {r EDA}
```

# 3. Model
*Describe your modeling approach and show how you arrived at your final
model.*

``` {r model}
```

# 4. Validation / Goodness of Fit

``` {r fit}
```

## 4.1 CV
*Validate your model with cross-validation and describe how your predictions
are useful (accuracy vs. generalizability).*

## 4.2 Map & Visualizations
*Provide additional maps and data visualizations to show that your model is
useful.*

# Conclusion
## Use Case Results
*Talk about how your analysis meets the use case you set out to address.*

## Improvements
*What could you do to make the analysis better?*