---
title: "MUSA508_FINAL_SFPARK"
date: "December 10, 2021"
editor_options: 
  chunk_output_type: console
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: "hide"
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
## Use Case & Motivation
*Motivate the analysis – “What is the use case; why would someone want to
replicate your analysis and why would they use this approach?”*

## Method
*What is the spatial or space/time process?*

``` {r intro}
```

# 1. Data
*Describe the data you used.*

``` {r data}
```

# 2. Exploratory Data Analysis
*Describe your exploratory analysis using maps and plots.*

``` {r EDA}
```

# 3. Model
*Describe your modeling approach and show how you arrived at your final
model.*

``` {r model}
```

# 4. Validation / Goodness of Fit

``` {r fit}
```

## 4.1 CV
*Validate your model with cross-validation and describe how your predictions
are useful (accuracy vs. generalizability).*

## 4.2 Map & Visualizations
*Provide additional maps and data visualizations to show that your model is
useful.*

# Conclusion
## Use Case Results
*Talk about how your analysis meets the use case you set out to address.*

## Improvements
*What could you do to make the analysis better?*